\myfont Implementation and evaluation of register allocation for translating LLVM-{}- to x86-64

### 2024-01-14 article.tex
blah

void dot(vector *a, vector *b) {
  a->x += b->x;
  a->y += b->y;     // -------->
}
     \end{minted}
     \caption{Subfield access}\label{fig:vector.c}
   \end{minipage}
   \begin{minipage}{0.48\textwidth}
     \centering
     \begin{minted}[linenos,bgcolor=LightGray]{llvm}
; ...
%6 = getelementptr %struct.vector,
            %i8* %1, i64 0, i32 2
%7 = load i32, i32* %6
%8 = getelementptr %struct.vector,
            %i8* %0, i64 0, i32 2
%9 = load i32, i32* %8
%10 = add i32 %9, %7
store i32 %10, i32* %8
; ...
     \end{minted}
     \caption{\texttt{clang -O1 -S -emit-llvm}}\label{fig:vector.ll}
   \end{minipage}
\end{figure}
\noindent Essentially \texttt{gep} instructions allow for unbounded array lookups, in that a \texttt{getelementptr} instruction is followed by the base address and one or more scaled offsets. What happens in Figure \ref{fig:vector.ll} is \texttt{vector} type is reduced to an \texttt{int} array, because it essentially is: in C, struct fields are generally  padded such that their respective memory addresses are aligned on four- or eight-byte boundaries \cite{cstd}.
As such, accessing the third field entry corresponds to accessing the third element of an \texttt{int} array. The \texttt{getelementptr} instruction gets the pointer of this element by multiplying the size with the index and adding that to the base address.
Therefore, the x86 that is generated essentially amounts to an \texttt{imul} instruction as can be seen in Figure

\begin{figure}[H]
    \centering
   \begin{minipage}{0.48\textwidth}
     \centering
     \begin{minted}[linenos,bgcolor=LightGray]{gas}
movq %rdi, %rcx
movq $0, %rax
imulq $0, %rax
movq $2, %rax
imulq $8, %rax
     \end{minted}
     \caption{\texttt{dune exec build -{}- -t x86}}\label{fig:square0s}
   \end{minipage}
\end{figure}

@@

\caption{\label{tab:dumb-fib}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1000}}

@@

@@

\noindent While it's not a complete 1:1 in terms of runtime as the Tiger compiler seems to perform slightly better than Greedy 0 (by \textasciitilde 30\% or so), that does make sense as there are marginally more instructions involved with each operation, not to mention the extra care taken to ensure calling conventions are upheld.

@@

Integer factorization is another interesting algorithm. Generally considered to be NP, here are the results of factorizing known primes of inreasing bit length. For \(n=1024\).

@@

16777213 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 167130& \cellcolor[rgb]{0.759765625,0.9990234375,0.75} 1.04160x& \cellcolor[rgb]{0.7509765625,0.9990234375,0.75} 1.00651x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05411x& \cellcolor[rgb]{0.7626953125,0.9990234375,0.75} 1.05372x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04686x& \cellcolor[rgb]{0.958984375,0.9990234375,0.75} 1.84213x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.33215x& \cellcolor[rgb]{0.9990234375,0.9873046875,0.75} 2.34511x\\
33554393 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 175409& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06219x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06070x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06415x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06212x& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06039x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.12199x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.78336x& \cellcolor[rgb]{0.9990234375,0.9736328125,0.75} 2.79956x\\
67108859 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 185779& \cellcolor[rgb]{0.7685546875,0.9990234375,0.75} 1.07588x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07923x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08373x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08043x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07895x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.49571x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38089x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.40121x\\
134217689 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 197302& \cellcolor[rgb]{0.7802734375,0.9990234375,0.75} 1.12173x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10225x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12120x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.12080x& \cellcolor[rgb]{0.779296875,0.9990234375,0.75} 1.11794x& \cellcolor[rgb]{0.9990234375,0.9677734375,0.75} 3.00248x& \cellcolor[rgb]{0.9990234375,0.9306640625,0.75} 4.17628x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.20718x\\
268435399 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 218313& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14571x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14655x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14531x& \cellcolor[rgb]{0.78515625,0.9990234375,0.75} 1.14376x& \cellcolor[rgb]{0.7861328125,0.9990234375,0.75} 1.14572x& \cellcolor[rgb]{0.9990234375,0.9501953125,0.75} 3.55232x& \cellcolor[rgb]{0.9990234375,0.9033203125,0.75} 5.06141x& \cellcolor[rgb]{0.9990234375,0.90234375,0.75} 5.08687x\\
536870909 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 245682& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18408x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18299x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18404x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18194x& \cellcolor[rgb]{0.794921875,0.9990234375,0.75} 1.18307x& \cellcolor[rgb]{0.9990234375,0.9296875,0.75} 4.20818x& \cellcolor[rgb]{0.9990234375,0.8720703125,0.75} 6.06747x& \cellcolor[rgb]{0.9990234375,0.8701171875,0.75} 6.13899x\\
1073741789 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 285913& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21667x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21894x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21755x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21668x& \cellcolor[rgb]{0.8037109375,0.9990234375,0.75} 1.21706x& \cellcolor[rgb]{0.9990234375,0.908203125,0.75} 4.89647x& \cellcolor[rgb]{0.9990234375,0.8369140625,0.75} 7.20172x& \cellcolor[rgb]{0.9990234375,0.8359375,0.75} 7.23917x\\

@@

\caption{\label{tab:factori32}Benchmark of \texttt{benches/factori32.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1000}}

>>>>

void dot(vector *a, vector *b) {
  a->x += b->x;
  a->y += b->y;     // -------->
}
     \end{minted}
     \caption{Subfield access}\label{fig:vector.c}
   \end{minipage}
   \begin{minipage}{0.48\textwidth}
     \centering
     \begin{minted}[linenos,bgcolor=LightGray]{llvm}
; ...
%6 = getelementptr %struct.vector,
            %i8* %1, i64 0, i32 2
%7 = load i32, i32* %6
%8 = getelementptr %struct.vector,
            %i8* %0, i64 0, i32 2
%9 = load i32, i32* %8
%10 = add i32 %9, %7
store i32 %10, i32* %8
; ...
     \end{minted}
     \caption{\texttt{clang -O1 -S -emit-llvm}}\label{fig:vector.ll}
   \end{minipage}
\end{figure}
\noindent Essentially \texttt{gep} instructions allow for unbounded array lookups, in that a \texttt{getelementptr} instruction is followed by the base address and one or more scaled offsets. What happens in Figure \ref{fig:vector.ll} is \texttt{vector} type is reduced to an \texttt{int} array, because it essentially is: in C, struct fields are generally  padded such that their respective memory addresses are aligned on four- or eight-byte boundaries \cite{cstd}.
As such, accessing the third field entry corresponds to accessing the third element of an \texttt{int} array. The \texttt{getelementptr} instruction gets the pointer of this element by multiplying the size with the index and adding that to the base address.
Therefore, the x86 that is generated essentially amounts to an \texttt{imul} instruction as can be seen in Figure \ref{fig:gep}.

\begin{figure}[H]
    \centering
   \begin{minipage}{0.48\textwidth}
     \centering
     \begin{minted}[linenos,bgcolor=LightGray]{gas}
; ...
movq %rdi, %rcx
movq $0, %rax
addq %rax, %rcx
imulq $8, %rax
movq $2, %rax
imulq $4, %rax
addq %rax, %rcx
; ...
     \end{minted}
     \caption{\texttt{dune exec build -{}- -t x86}}\label{fig:gep}
   \end{minipage}
\end{figure}

\noindent where the base register  \texttt{\%1} has assignment \(\mathit{assign}[\texttt{\%1}] = \texttt{\%rdi}\) is moved into the scratch register \texttt{\%rcx}, and each of the offset operands are 1) moved into the other scratch register \texttt{\%rax}, 2) multiplied by the size of their respective types, and 3) added to the base register.

Unfortunately, the gep wasn't implemented properly in time for handin. Instead the code appears to be left over from what was jumbled together in the Tiger days, so it does work in \textit{some} cases but for from all.

@@

\caption{\label{tab:dumb-fib}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1024}}

@@

\caption{\label{tab:dumb-fib-greedy}Benchmark of \texttt{benches/fib.ll} output by \mintinline{bash}{dune exec bench -- -f fib -n 1024}}

@@

\noindent While it's not a complete one to one in terms of runtime as the Tiger compiler seems to perform slightly better than Greedy 0 (by \textasciitilde 30\% or so), that does make sense as there are marginally more instructions involved with each operation, not to mention the extra care taken to ensure calling conventions are upheld.

@@

\subsubsection{\lstinline!benches/loopn.ll!}

Another simple example is a program that only consists of a loop running for \(n\) iterations. There stem from the \texttt{tests/loopn.c} file but with the call to \texttt{printf} removed such that no unnecessary time is spent doing I/O. Depending on the level of optimization chosen, either the mutation strategy of stack allocated variables or phi nodes are used.  For the case of using the stack, the simplicty of such a program shines through as can be seen on Table \ref{tab:loopn0}:

\begin{table}[H]
\makebox[\textwidth][c]{
\begin{NiceTabular}{|c|ccccccccc|}
\hline
arg(s) & clang & simple 12 & simple 2 & briggs 12 & briggs 2 & linear 12 & linear 2 & greedy 12 & greedy 0 \\
\hline
1048575 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 2238539& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.56041x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63938x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.56101x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63915x& \cellcolor[rgb]{0.9990234375,0.9814453125,0.75} 2.56111x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 19.26926x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 27.70065x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.60162x\\
2097151 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 4333585& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.61339x& \cellcolor[rgb]{0.9990234375,0.9765625,0.75} 2.69416x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.61334x& \cellcolor[rgb]{0.9990234375,0.9765625,0.75} 2.69427x& \cellcolor[rgb]{0.9990234375,0.9794921875,0.75} 2.61352x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 19.87471x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 28.58590x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.78097x\\
4194303 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 8529189& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63937x& \cellcolor[rgb]{0.9990234375,0.9765625,0.75} 2.72120x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63934x& \cellcolor[rgb]{0.9990234375,0.9765625,0.75} 2.72124x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.63938x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 20.17961x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.03173x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37560x\\
8388607 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 16919352& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.65269x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73526x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.65275x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.73534x& \cellcolor[rgb]{0.9990234375,0.978515625,0.75} 2.65264x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 20.33697x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.26211x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68258x\\
16777215 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 33697415& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.65966x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74260x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.65966x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74263x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.65964x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 20.41817x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.39601x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.84065x\\
33554431 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 67254339& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66311x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74627x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66312x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74628x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66311x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 20.45872x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.46374x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.91968x\\
67108863 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 134365449& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66491x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74816x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66488x& \cellcolor[rgb]{0.9990234375,0.9755859375,0.75} 2.74816x& \cellcolor[rgb]{0.9990234375,0.9775390625,0.75} 2.66492x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 20.47947x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.49845x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96011x\\
\hline
\end{NiceTabular}}
\caption{\label{tab:loopn0}Benchmark of \texttt{benches/factori64.ll} output by \mintinline{bash}{dune exec bench -- -f factori32 -n 1024}}
\end{table}

\noindent Despite having no optimizations enabled and using the stack to store the counter, \texttt{clang} manages to build an executable wherein the loop consists of no more than 7 instructions (see Figure \ref{fig:clang-loop}). In comparison, mine is 19 (see Figure \ref{fig:my-loop0}):

\begin{figure}[H]
     \centering
     \begin{minipage}[b]{0.32\textwidth}
\begin{minted}[bgcolor=LightGray]{gas}
# ...
movslq 0xc(%rsp),%rax
cmp    0x10(%rsp),%rax
jae    0x1199 <main+89>
mov    0xc(%rsp),%eax
add    $0x1,%eax
mov    %eax,0xc(%rsp)
jmp    0x1179 <main+57>
# ...
     \end{minted}
     \caption{TODO}\label{fig:clang-loop}
   \end{minipage}
   \begin{minipage}[b]{0.65\textwidth}
   \begin{minipage}[b]{0.50\textwidth}
  \begin{minted}[bgcolor=LightGray]{gas}
# ...
movq	%rsi, %rax
movl	(%rax), %eax
movq	$0, %rdx
movl	%eax, %edx
movq	%rdx, %rax
movq	$0, %rbx
movl	%eax, %ebx
movq	%rdi, %rax
movq	(%rax), %rax
movq	%rax, %rdx
     \end{minted}
   \end{minipage}
   \begin{minipage}[b]{0.50\textwidth}
  \begin{minted}[bgcolor=LightGray]{gas}
movq	%rax, %rdx
movq	%rbx, %rax
movq	%rdx, %rcx
cmpq	%rcx, %rax
movq	$0, %rdx
setb	%dl
movq	%rdx, %rax
cmpq	$0, %rax
je	_main$20
jmp	_main$17
# ...
     \end{minted}
   \end{minipage}
     \caption{TODO}\label{fig:my-loop0}
   \end{minipage}
\end{figure}
\noindent It does make sense that my compiler nearly takes 3x as many cycles considering that nearly 3x as many instructions are being executed.   Because od the way values are stored, there is literally no benefit from Briggs coalescing since no phi nodes are involved ever.
I was surprised to see Greedy and Linear 2 to be that much slower, but on second thought it makes perfect sense since loading variables from the stack before every single arithmetic/logic operation is considerably slower. What's more surprising is the fact that Tiger is nearly 4x faster than any Greedy (see Table \ref{tab:loopn0.tig}).
It's presumably just because all binary operations consist of  about three instructions or so.

\begin{table}[H]
\makebox[\textwidth][c]{
\begin{NiceTabular}{|c|ccccccccc|}
\hline
arg(s) & clang & greedy 12 & greedy 8 & greedy 6 & greedy 4 & greedy 2 & greedy 1 & greedy 0 & tiger \\
\hline
1048575 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 2237215& \cellcolor[rgb]{0.9990234375,0.75,0.75} 27.71735x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.15447x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.62296x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.62342x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.62283x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.62360x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 36.62324x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.97018x\\
2097151 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 4334414& \cellcolor[rgb]{0.9990234375,0.75,0.75} 28.58063x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.28996x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.77377x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.77391x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.77373x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.77353x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.77356x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.29302x\\
4194303 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 8529224& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.03182x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 37.88371x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37535x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37541x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37546x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37544x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.37543x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.46122x\\
8388607 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 16919774& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.26144x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.18576x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68170x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68169x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68175x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68169x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.68163x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.54653x\\
16777215 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 33698040& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.39364x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.34210x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.83996x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.83997x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.83991x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.83995x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.83993x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.59127x\\
33554431 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 67253512& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.46538x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.42125x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.92015x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.92019x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.92020x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.92016x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.92020x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.61419x\\
67108863 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 134364289& \cellcolor[rgb]{0.9990234375,0.75,0.75} 29.49788x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.46111x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96041x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96050x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96052x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96043x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 38.96040x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.62584x\\
\hline
\end{NiceTabular}}
\caption{\label{tab:loopn0.tig}Benchmark of \texttt{benches/loopn0.ll} output by \mintinline{bash}{dune exec bench -- -f loopn0 -n 64}}
\end{table}

\noindent Even with \texttt{-O1} where phi nodes are introduced, does the Briggs allocator seem to affect anything as can be seen in Table \ref{tab:loopn1}. This is because the only two variables eligible for coalescing, \texttt{\%9} and \texttt{\%10}, are in conflict as both are used at separate points.

@@

131071 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 401485& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65252x& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65413x& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65379x& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65354x& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65376x& \cellcolor[rgb]{0.912109375,0.9990234375,0.75} 1.65265x& \cellcolor[rgb]{0.908203125,0.9990234375,0.75} 1.63694x& \cellcolor[rgb]{0.9990234375,0.8056640625,0.75} 8.19401x\\
262143 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 663679& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.79023x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.79079x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.78950x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.79092x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.78914x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.78887x& \cellcolor[rgb]{0.9462890625,0.9990234375,0.75} 1.79096x& \cellcolor[rgb]{0.9990234375,0.7587890625,0.75} 9.69501x\\
524287 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 1188029& \cellcolor[rgb]{0.9697265625,0.9990234375,0.75} 1.88330x& \cellcolor[rgb]{0.9697265625,0.9990234375,0.75} 1.88252x& \cellcolor[rgb]{0.96875,0.9990234375,0.75} 1.88192x& \cellcolor[rgb]{0.9697265625,0.9990234375,0.75} 1.88240x& \cellcolor[rgb]{0.9697265625,0.9990234375,0.75} 1.88242x& \cellcolor[rgb]{0.96875,0.9990234375,0.75} 1.88211x& \cellcolor[rgb]{0.9697265625,0.9990234375,0.75} 1.88284x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 10.79175x\\
1048575 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 2237256& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93727x& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93740x& \cellcolor[rgb]{0.982421875,0.9990234375,0.75} 1.93674x& \cellcolor[rgb]{0.982421875,0.9990234375,0.75} 1.93715x& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93728x& \cellcolor[rgb]{0.9833984375,0.9990234375,0.75} 1.93739x& \cellcolor[rgb]{0.982421875,0.9990234375,0.75} 1.93690x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.42118x\\
2097151 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 4334546& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96774x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96724x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96767x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96746x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96749x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96756x& \cellcolor[rgb]{0.990234375,0.9990234375,0.75} 1.96753x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 11.81912x\\
4194303 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 8529268& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98350x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98348x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98353x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98339x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98339x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98358x& \cellcolor[rgb]{0.994140625,0.9990234375,0.75} 1.98355x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 12.04037x\\
8388607 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 16918996& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99159x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99166x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99166x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99160x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99166x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99165x& \cellcolor[rgb]{0.99609375,0.9990234375,0.75} 1.99162x& \cellcolor[rgb]{0.9990234375,0.75,0.75} 12.13501x\\

@@

\caption{\label{tab:loopn1}Benchmark of \texttt{benches/loopn1.ll} output by \mintinline{bash}{dune exec bench -- -f loopn1 -n 64}}

\subsubsection{\lstinline!benches/phis.ll!}
To actually see the benefit of coalescing in action I have introduced some generated tests of \texttt{benches/phis\{0-7\}.ll}. Generated with \texttt{benches/phis.py <n> <m>} that produces a \texttt{.ll} file that consists of a loop hardcoded for \(n\) repetitions with \(m\) phi nodes.

Integer factorization is another interesting algorithm. Generally considered to be NP, here are the results of factorizing known primes of inreasing bit length:
16777213 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 168400& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04991x& \cellcolor[rgb]{0.7568359375,0.9990234375,0.75} 1.02776x& \cellcolor[rgb]{0.76171875,0.9990234375,0.75} 1.04829x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04435x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04465x& \cellcolor[rgb]{0.9560546875,0.9990234375,0.75} 1.82795x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.31432x& \cellcolor[rgb]{0.9990234375,0.98828125,0.75} 2.32746x\\
33554393 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 175723& \cellcolor[rgb]{0.7646484375,0.9990234375,0.75} 1.06090x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03777x& \cellcolor[rgb]{0.765625,0.9990234375,0.75} 1.06480x& \cellcolor[rgb]{0.7607421875,0.9990234375,0.75} 1.04465x& \cellcolor[rgb]{0.7587890625,0.9990234375,0.75} 1.03702x& \cellcolor[rgb]{0.9990234375,0.9951171875,0.75} 2.12225x& \cellcolor[rgb]{0.9990234375,0.974609375,0.75} 2.78015x& \cellcolor[rgb]{0.9990234375,0.9736328125,0.75} 2.79323x\\
67108859 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 185913& \cellcolor[rgb]{0.763671875,0.9990234375,0.75} 1.05642x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.08067x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08352x& \cellcolor[rgb]{0.76953125,0.9990234375,0.75} 1.07946x& \cellcolor[rgb]{0.7705078125,0.9990234375,0.75} 1.08349x& \cellcolor[rgb]{0.9990234375,0.9833984375,0.75} 2.50046x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.38055x& \cellcolor[rgb]{0.9990234375,0.955078125,0.75} 3.39737x\\
134217689 & \cellcolor[rgb]{0.75,0.9990234375,0.75} 199433& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11084x& \cellcolor[rgb]{0.7783203125,0.9990234375,0.75} 1.11393x& \cellcolor[rgb]{0.771484375,0.9990234375,0.75} 1.08952x& \cellcolor[rgb]{0.77734375,0.9990234375,0.75} 1.11192x& \cellcolor[rgb]{0.775390625,0.9990234375,0.75} 1.10441x& \cellcolor[rgb]{0.9990234375,0.96875,0.75} 2.97084x& \cellcolor[rgb]{0.9990234375,0.931640625,0.75} 4.13960x& \cellcolor[rgb]{0.9990234375,0.9326171875,0.75} 4.12783x\\
